The Heavy Photon Search Test Detector

### 2014-10-24 reviewer_comments_answers.txt

v5.0
Reviewer #1
The paper is technically correct and well done concerning the presentation of the physics program, the instrumental realization and the first experience. However, the publication should provide at the end not only a report on the work which has been done but also contain results of common interest. In order to be more specific, the introduction does not give exact values for the required resolutions with respect to tracking and timing or in particular the energy resolution of the calorimeter.

ANSWER
The revised version contains some significant changes and improvements and spells out the HPS requirements more clearly in the introduction. We hope this will clarify what the HPS Test Run performance limitations were.

In particular we better specified the requirements of both the SVT and ECal, both in the introduction and also in the sections describing the two detectors and what was driving their design (Sec. 4 and 5). Similarly,  we have better related the Test Run performance and its limitations to both the immediate needs of the Test Run (to demonstrate functionality and measure trigger rates) and the ultimate HPS requirements. For example we include the expected resolution of tracks and clusters in the ECal, but also explain what was adequate for the Test Run trigger and what needs improving for HPS offline reconstruction.

There is also no comment on the expected dynamic range to judge the acievable or achieved energy resolution. In most cases there is only the comment that the performance is sufficient or can be improved.

ANSWER
In the case of the SVT we list the 3 mip dynamic range with linear gain of the APV25 chip (Sec. 7.1.1). We also quote the typical noise and signal levels to get a feeling for the S/N~25. The resolution of the ADC is 14-bits which is discussed in Sec. 6.2.

For the ECal the maximum deposited energy in a single crystal is 4.2GeV with a beam energy of 6.6GeV. The dynamic range was checked for the CLAS IC (which was refurbished for HPS Test Run) and was found to be linear across this range to within 1-2%. The 12-bit resolution of the FADC means that we will get a resolution of about 1ADC/MeV which is enough given that our noise is of the order of 10MeV.

We improved the description of the readout chain for the ECal and in particular added comments about the requirements to be able to capture the maximum expected deposited energy and also the resolution of the ADC. We hope that this clarifies the dynamic range and what is needed for HPS.

In particular, in case of the calorimeter section the achieved performance does not appear acceptable. The high noise level as well as the extremely high thresholds indicate that there is a significant improvement necessary. With such a high threshold a shower reconstruction appears impossible except one focuses on the highest energies only. Even the detected multiplicities a without meaning. Therefore, the authors should focus more consequent on a clear picture of the overall requirements, the achievements, any general limitations and the options to reach the final goals. In that case the complex work would become more valuable for an external reader.

ANSWER
Parts of the test run performance are indeed undesirable for the HPS experiment; the dead channel count, noise level, and energy resolution were all substandard. Despite those facts, the Ecal performance was acceptable to provide a trigger for the Test Run and reliably measure normalized trigger rates.

The number of non-functioning channels (reasons listed in the paper) is much less than desirable for triggering efficiently on A' in HPS. It has already been improved by installing bigger APDS, modified amplifiers, and new HV distribution for all the crystals in the detector. The new readout chain (new APDs and preamplifiers) show noise levels of about 7.5MeV bringing thresholds down to closer to 10MeV. Simulations show that with a wel- calibrated detector the Ecal esolution will about 3.6% in HPS.

The A' trigger is relatively insensitive to the energy resolution as most of the rejection comes from combined kinematic cuts and time coincidence. Even the simple calibration procedures applied in the Test Run and the relatively high noise in the system were sufficient for Test Run triggering. The 4.5% energy resolution goal was predicated not on trigger-ability, but the desire to use the Ecal to improve on the SVT's measurement of the momentum of electrons and positrons, and so improve on the HPS invariant mass resolution.

We have edited the paper pointing out the requirements of the ECal in the Sec. 5 and added further explanations in the ECal performance section 7.2.

Reviewer #2
The paper describes in considerable technical detail a test experiment for the Heavy Photon Search (HPS) at Jefferson Lab.
The paper is well-written, comprehensive, and gives a very detailed technical description of the apparatus, results, and
prospects for the data taking run.  I recommend that it be published with the following suggestion:

The details of the HPS experiment are not provided clearly.  From other sources , I understand that HPS proposes to scatter
an electron beam from a tungsten target and that the A' would be generated in the scattering.  The test run here used a photon
beam generated from conversion in a gold foil. I suggest that the process to be measured in the HPS experiment (projectile and target) 
be clearly specified at the beginning of the paper.

ANSWER
We have updated the introduction with more information to better explain the HPS experiment, and the differences between the Test Run and HPS. Hopefully this also clarifies the purpose of the Test Run and the relevance of its performance to HPS.

### 2014-10-24 hps-testrun-nim.tex
Comments from Tim

% \linenumbers

@@

The design of the SVT is primarily driven by physics requirements and constraints from the 
environment at the interaction region. 
The \Aprime{} decay products have momenta in the range of 0.4-2.0~GeV$/$c (for a 2.2~GeV beam), so 
multiple scattering 
dominates mass and vertexing uncertainties for any possible material budget. The SVT must therefore 
minimize the amount of material in the tracking volume. 
The signal yield for long-lived \Aprime{} is very small, so 
the rejection of prompt vertices must be exceedingly pure, on the order of $10^{-7}$, in order to 
eliminate all prompt backgrounds. To achieve the required vertexing performance the first layer of the 
SVT must be placed no more than about 10~cm downstream of the target. At that distance, it is found 
that the active region of a sensor can be placed as close as 1.5~mm from the center of the beam, 
defining the 15~mrad ``dead zone'' mentioned previously, to maximize low-mass \Aprime{} acceptance 
with decay products nearly collinear with the beam axis. At the edge of this ``dead zone'', the 
radiation dose approaches $10^{15}$~electrons/cm$^2$/month, or roughly 
$3 \times 10^{14}$~\fluenceunit{}/month~\cite{Rashevskaya:2002nd}, 
requiring the sensors to be actively cooled. 
Meanwhile, very low-energy delta rays from 
beam-gas interactions would multiply the density of background hits, so the SVT must operate inside the 
beam vacuum.  Finally, in order to protect the sensors, the detector must be movable so that it can be 
retracted during periods of uncertain beam conditions or beam tuning.

@@

High background occupancies, up to 4~MHz$/$mm$^{2}$ locally,  in the region closest to the beam result 
from beam electrons undergoing multiple scattering in the target. These background hits are rejected by  
requiring the time reconstruction of each hit to be better than 2~ns.

@@

%\subsubsection{Front-end Electronics}
Despite having only small spots with very high occupancy (up to 4~MHz/mm$^2$) closest to the primary 
beam, the rates are still high and lowering the peak occupancy to 
approximately 1\% for tracking requires a trigger window and hit time tagging of roughly 8~ns. The 
ECal readout and trigger described in Sec.~\ref{sec:fadc} can achieve such resolution. To reach this 
performance the sensors for the SVT are readout by the APV25 ASIC developed for the CMS 
experiment at CERN~\cite{French:2001xb}. The APV25 ASIC can capture successive samples of the shaper output 
in groups of three at a sampling rate of approximately 40~MHz.  By fitting the known 
$CR$-$RC$ shaping curve to these samples, the initial time of the hit can be determined to a precision 
of 2~ns for S/N$\approx$25~\cite{Friedl:2009zz}.  For electron beam running, six-sample readout and 
the shortest possible shaping time (35~ns) is used to best distinguish hits that overlap in time.
The APV25 ASICs are hosted on simple FR4 hybrid readout boards, outside the tracking volume, with a 
short twisted-pair pigtail cable to provide power and configuration and signal readout. Along with a 
single sensor, these are glued to a polyimide-laminated carbon fiber composite backing making 
up a half-module. A window is machined in the carbon fiber leaving only a frame around the periphery 
of the silicon to minimize material. A 50~$\mu$m sheet of polyimide is laminated to the surface of the 
carbon fiber with 1~mm overhang at all openings to ensure good isolation between the back side of the 
sensor, carrying high-voltage bias, and the carbon fiber which is held near ground.

@@

The electromagnetic calorimeter (ECal), installed downstream of the PS dipole magnet, performs two 
essential functions for the experiment: it provides a trigger signal to select what events to read out from 
the detector sub-systems and is used to identify electrons and positrons. 
The technology and design choices are largely driven by the need for a compact forward design 
covering the SVT \Aprime{} 
acceptance and able to measure the energy and positions of electrons and positrons with energy between
$0.5$ and $\approx$4~GeV. It needs granularity and signal 
readout speed to handle 1~MHz/cm$^{2}$ of electromagnetic background and radiation hardness. 
Even modest energy resolution is adequate for triggering. HPS requires better 
energy resolution, $\sigma(E)/E<5\%/\sqrt{E}$, so that the ECal energy measurement can be used in 
combination with that from the SVT to improve the overall momentum resolution.

@@

\subsection{Trigger System}
\label{sec:trigger}
The trigger system is designed to select time coincidences of electromagnetic clusters in the top and bottom 
halves of the ECal which meets kinematic conditions satisfied by \Aprime{} decays and minimize backgrounds.
The trigger system needs to be essentially dead-time free, handle rates up to 50~kHz, and supply a trigger 
signal which jitters $<8$~ns from the actual event time in order to minimize backgrounds from out-of-time hits in the SVT. 
Figure~\ref{fig:hps_trigger_cal} shows a schematic overview of each stage of the system. 
 \begin{figure}[b]
\begin{center}
 \includegraphics[width=8cm]{hps_trigger_cal}
\caption{\small Block diagram of the ECAL trigger system consisting of the FADC that samples and digitizes 
signals for each detector channel and sends them for cluster finding in the CTP. The CTP clusters are 
sent to the SSP where the final trigger decision is taken based on pairs of clusters in both halves of the 
ECal. The decision is sent to the Trigger Supervisor (TS) that generates the necessary signals to read out 
the sub-detectors.
 \label{fig:hps_trigger_cal}}
\end{center}
 \end{figure}
Each channel on the FADC board has an independent data path to send 5-bit pulse energy and 3-bit 
pulse arrival time information every 32~ns to a trigger processing board (CTP), which is in the same 
crate. The 3-bit pulse arrival time allows the trigger to know the pulse timing at 4~ns resolution. 
Contrary to the readout path 
described in Sec.~\ref{sec:fadc}, this energy is a pedestal-subtracted time-over-threshold sum with 
programmable offsets and minimum threshold discriminator for each channel. With input from all 
FADC channels, i.e. one half of the ECal, the CTP performs cluster finding and calculates cluster 
energy and timing information. The 3x3 fixed-window, highly parallel, FPGA-based cluster 
algorithm simultaneously searches for up to 125 clusters with energy sum larger than 
the programmable energy threshold set to about 270~MeV. This high threshold didn't hurt the trigger rate 
rate studies for the HPS Test run since only clusters with high energies were studied but for HPS this 
threshold will need to be lower.  Crystals in the 
fixed-window are included in the sum if the leading edge of the pulse occurred within a 32~ns time 
window to take into account clock skew and jitter throughout the system.
The CTP only accepts clusters with the locally highest energy 3x3 window to deal with overlapping and 
very large clusters. The sub-system board (SSP) receives the clusters from the top and bottom half CTP 
at a maximum rate of 250~MHz and searches for pairs of clusters in an 8~ns wide coincidence window. The 
SSP sends triggers to the trigger supervisor (TS), which generates all the necessary signals and 
controls the entire DAQ system readout through the trigger interface units installed in every crate that 
participate in the readout process.

@@

During the HPS Test Run, for the most part the trigger system 
required only a single cluster in either the top or bottom ECal halves. It was tested to trigger rates 
above 100~kHz by lowering thresholds.

@@

The SVT DAQ is based on the Reconfigurable Cluster Element (RCE) and cluster 
interconnect concept developed at SLAC as generic building blocks for DAQ systems. 
The RCE is a generic computational building block, housed on a separate daughter card called 
Data Processing Module (DPM), that is realized on an ATCA front board called the Cluster On Board 
(COB), see Figure~\ref{fig:cob}.
 \begin{figure}[]
\begin{center}
\includegraphics[width=6cm]{svt_daq_module}
\caption{\small The SVT DAQ COB board with four data processing daughter cards (DPMs) visible on the left side.
\label{fig:cob}}
\end{center}
\end{figure}
The first generation RCE used in the HPS Test Run consisted of a Virtex~5 FPGA with 1~GB of DDR3 RAM. 
A schematic overview of the system is shown in Figure~\ref{fig:svtdaq}. 
 \begin{figure}[]
\begin{center}
\includegraphics[width=8cm]{svt-daq-sketch}
\caption{\small Block diagram overview of the SVT DAQ.
\label{fig:svtdaq}}
\end{center}
\end{figure}
The analog outputs of up to 12 SVT half-modules (60 APV25 ASICs) are digitized on the 
Rear-Transition-Module (RTM), a custom board on the back side of the ATCA crate, interfacing the 
HPS-specific readout to the generic DAQ components on the COB. A preamplifier converts the APV25 ASIC 
differential current output to a different voltage output scaled to the sensitive range of a 14-bit ADC 
operating at the system clock of 41.667~MHz. The RTM is organized 
into four sections with each section supporting three SVT half-module hybrids (15 APV25 ASICs). The 
RTM also includes a 4-channel fiber-optic module and supporting logic which is used to interface 
to the JLab trigger system supervisor. Each section of the RTM is input to a DPM which apply 
thresholds for data reduction and organizes the sample data into UDP datagrams. The DPM also hosts 
an I$^{2}$C controller used to configure and monitor the APV25 ASICs. A single ATCA crate with two 
COB cards was used, one supporting four DPMs and one supporting three DPMs and one DPM that is 
configured as the trigger and data transmission module. The two COB cards and their DPMs are 
interconnected with a 10~Gb/s switch card~\cite{Larsen:2011zb} which also hosts two 1~Gb/s Ethernet 
interfaces to the external SVT DAQ PC.

@@

In order to minimize cable length for the analog APV25 ASIC output signal the ATCA crate was located 
approximately 1~m from the beam line, next to our cable vacuum feed-throughs.   
Before shielding with lead-blankets and borated polyethylene was arranged, we observed two failures of normally reliable ATCA 
crate power supplies, time-correlated to beam instabilities.

@@

After a trigger is received, the amplitude of every APV25 ASIC is sampled and digitized in the six consecutive time 
bins associated with the trigger time. A data reduction algorithm is applied 
requiring three out of six samples to be above two times the noise level and that the third sample is 
larger than the second or that the fourth sample is larger than the third.
The typical, pedestal subtracted, pulse shape obtained is shown in Figure~\ref{fig:pulse_shape}. 
As the figure demonstrates,  the SVT was well timed-in to the trigger with the rise of the pulse at the 3rd 
sampling point. 
\begin{figure}[]
\begin{center}
\includegraphics[width=7cm]{run1351_110513_samples_L1_top.pdf}
\caption{\small Accumulation of six pedestal-subtracted samples from individual SVT channels associated 
with hits on tracks.
\label{fig:pulse_shape}}
\end{center}
\end{figure}
In order to find the time, $t_0$, and amplitude of each hit, the six samples from each channel are fitted 
to an ideal $CR-RC$ function. Note that in the HPS Test Run the APV25 ASICs were operating with a 50~ns 
shaping time. These hits are passed through a simple clustering algorithm which forms clusters by 
grouping adjacent strips with the position of a cluster on the sensor determined by the 
amplitude-weighted mean.
With a linear gain up to $\approx 3$~MIPs, the cluster charge for hits associated with a track follow 
the characteristic Landau shape, see Figure~\ref{fig:cluster_pulse}.  
\begin{figure}[]
\begin{center}
\includegraphics[width=7cm]{mip_top_layer_2.png}
\caption{\small The cluster charge distribution for hits associated with a track follow the characteristic Landau shape. 
\label{fig:cluster_pulse}}
\end{center}
\end{figure}
A noise level between $1.1-1.5\times 10^{3}$ electrons was established through multiple calibration 
runs giving a $S/N$ of $21-25$, in line with the requirement for HPS. 
Radioactive source tests were used to 
provide the absolute charge normalization.
After clustering hits on a sensor, the hit time for each cluster is computed 
as the amplitude-weighted average of the individually fitted $t_0$ on each channel. The 
$t_0$ resolution is studied by comparing the cluster hit time with the average of all cluster hit times on 
the track shown in Figure~\ref{fig:tracktime}. 
After correcting for offsets from each sensor (time-of-flight and clock phase) and accounting for the 
correlation between the $t_0$ and track time,  the extracted $t_0$ resolution is 2.6~ns. This is 
somewhat worse than the approximately 2~ns resolution expected for S/N$=$25 which we attribute to the 
true pulse shape differing from our idealized fit function which will be improved in the 
future~\cite{Friedl2007385}. Reducing 
the APV25 ASIC pulse shaping time to 35~ns will also improve time resolution. 
These results show that we can operate with the six sample readout mode of the APV25 ASIC and 
achieve time resolution adequate for pileup rejection during electron running in HPS. 
\begin{figure}[]
\begin{center}
\includegraphics[width=7cm]{run1351_time_residual_top_layer2.pdf}
\caption{\small The residual of individual cluster times with the average of all clusters on the track. 
\label{fig:tracktime}}
\end{center}
\end{figure}

@@

\subsection{Trigger Rate Comparisons}
Trigger rates observed in the HPS Test Run are dominated by \ee{} pairs produced in the 
converter. In simulated events, the rate of triggers depend on the modeling of the pairs' angular 
distribution and the subsequent multiple Coulomb scattering in the converter. Rates from different 
converter 
thicknesses are used to study the varying multiple Coulomb scattering contribution (pair production 
angular distribution is constant), and are compared with Geant4~\cite{Agostinelli2003250} standard multiple 
scattering model and EGS5~\cite{egs5}. 
%\begin{center}
\begin{table}
{\small
\begin{tabular}{|l|c|c|c|}
\hline
\bf Converter (\%~$X_0$) & \bf 1.60 & \bf 0.45 &	\bf 0.18 \\
\hline
EGS5 &	1162 $\pm$ 112 &	255 $\pm$ 28 &	94 $\pm$ 17	\\
\hline
Geant4 & 2633 $\pm$ 250 & 	371 $\pm$ 38 &	114 $\pm$ 18 \\
\hline
Observed 	& 1064 $\pm$ 2 & 196 $\pm$ 1 &	92 $\pm$ 1 \\						
\hline
\end{tabular}
\caption{ Observed and predicted event rate (in Hz) normalized to 90~nA for three different converter 
thicknesses. The uncertainty on the prediction includes systematic uncertainties from ECal alignment, background normalization, beam current normalization and limited statistics in the simulation.
\label{results}}
}
\end{table}
%\end{center}
Restricting to a well calibrated region of the ECal and to clusters with energy above the trigger turn-on, 
we see agreement with the rates predicted by the \egs{} simulation program after subtracting the 
``no converter'' background , see Table~\ref{results}.
This gives further confidence that the dominant source of background occupancy for HPS, multiple 
Coulomb scattered beam electrons, is well 
described.
%~\cite{Attwood:2005zz,Shen:1978ha,Gottschalk1993467} .

>>>>

 %\linenumbers

@@

The design of the SVT is primarily driven by physics requirements and constraints from the 
environment at the interaction region. 
The \Aprime{} decay products have momenta in the range of 0.4-2.0~GeV$/$c (for a 2.2~GeV beam), so 
multiple scattering 
dominates mass and vertexing uncertainties for any possible material budget. The SVT must therefore 
minimize the amount of material in the tracking volume. 
The signal yield for long-lived \Aprime{} is very small, so 
the rejection of prompt vertices must be exceedingly pure, on the order of $10^{-7}$, in order to 
eliminate all prompt backgrounds. To achieve the required vertexing performance the first layer of the 
SVT must be placed no more than about 10~cm downstream of the target. At that distance, it is found 
that the active region of a sensor can be placed as close as 1.5~mm from the center of the beam, 
defining the 15~mrad ``dead zone'' mentioned previously, to maximize low-mass \Aprime{} acceptance 
with decay products nearly collinear with the beam axis. At the edge of this ``dead zone'', the 
radiation dose approaches $10^{15}$~electrons/cm$^2$/month, or roughly 
$3 \times 10^{13}$~\fluenceunit{}/month~\cite{Rashevskaya:2002nd}, 
requiring the sensors to be actively cooled. 
Meanwhile, very low-energy delta rays from 
beam-gas interactions would multiply the density of background hits, so the SVT must operate inside the 
beam vacuum.  Finally, in order to protect the sensors, the detector must be movable so that it can be 
retracted during periods of uncertain beam conditions or beam tuning.

@@

High background occupancies, up to 4~MHz$/$mm$^{2}$ locally,  in the region closest to the beam result 
from beam electrons undergoing multiple scattering in the target. These background hits are rejected by  
requiring reconstruction of the hit time relative to the trigger with 2~ns resolution.

@@

%\subsubsection{Front-end Electronics}
Despite having only small spots with very high occupancy (up to 4~MHz/mm$^2$) closest to the primary 
beam, the rates are still high and lowering the peak occupancy to 
approximately 1\% for tracking requires a trigger window and hit time tagging of roughly 8~ns. The 
ECal readout and trigger described in Sec.~\ref{sec:fadc} can achieve such resolution. To reach this 
performance the sensors for the SVT are readout by the APV25 ASIC developed for the CMS 
experiment at CERN~\cite{French:2001xb}. The APV25 ASIC can capture successive samples of the shaper output 
in groups of three at a sampling rate of approximately 40~MHz.  By fitting the known 
$CR$-$RC$ shaping curve to these samples, the initial time of the hit can be determined to a precision 
of 2~ns for S/N$\approx$25~\cite{Friedl:2009zz}.  For electron beam running, six-sample readout and 
the shortest possible shaping time (35~ns) are used to best distinguish hits that overlap in time.
The APV25 ASICs are hosted on simple FR4 hybrid readout boards outside the tracking volume with a 
short twisted-pair pigtail cable to provide power, configuration, and signal readout. Along with a 
single sensor, these are glued to a polyimide-laminated carbon fiber composite backing making 
up a half-module. A window is machined in the carbon fiber leaving only a frame around the periphery 
of the silicon to minimize material. A 50~$\mu$m sheet of polyimide is laminated to the surface of the 
carbon fiber with 1~mm overhang at all openings to ensure good isolation between the back side of the 
sensor, carrying high-voltage bias, and the carbon fiber which is held near ground.

@@

The electromagnetic calorimeter (ECal), installed downstream of the PS dipole magnet, performs two 
essential functions for the experiment: it provides a trigger signal to select what events to read out from 
the detector sub-systems and is used to identify electrons and positrons. 
The technology and design choices are largely driven by the need for a compact forward design 
covering the SVT \Aprime{} 
acceptance and the ability to measure the energy and positions of electrons and positrons with energy between
0.5 and 6.6~GeV. It needs granularity and signal 
readout speed to handle 1~MHz/cm$^{2}$ of electromagnetic background as well as good radiation hardness. 
Even modest energy resolution is adequate for triggering. HPS requires better 
energy resolution, $\sigma(E)/E<5\%/\sqrt{E}$, so that the ECal energy measurement can be used in 
combination with that from the SVT to improve the overall momentum resolution.

@@

\subsection{Trigger System}
\label{sec:trigger}
The trigger system is designed to select time coincidences of electromagnetic clusters in the top and bottom 
halves of the ECal which meets kinematic conditions satisfied by \Aprime{} decays and minimize backgrounds.
The trigger system needs to be essentially dead-time free, handle rates up to 50~kHz, and supply a trigger 
signal which jitters $<8$~ns from the actual event time in order to minimize backgrounds from out-of-time hits in the SVT. 
Figure~\ref{fig:hps_trigger_cal} shows a schematic overview of each stage of the system. 
 \begin{figure}[b]
\begin{center}
 \includegraphics[width=8cm]{hps_trigger_cal}
\caption{\small Block diagram of the ECAL trigger system consisting of the FADC that samples and digitizes 
signals for each detector channel and sends them for cluster finding in the CTP. The CTP clusters are 
sent to the SSP where the final trigger decision is taken based on pairs of clusters in both halves of the 
ECal. The decision is sent to the Trigger Supervisor (TS) that generates the necessary signals to read out 
the sub-detectors.
 \label{fig:hps_trigger_cal}}
\end{center}
 \end{figure}
Each channel on the FADC board has an independent data path to send 5-bit pulse energy and 3-bit 
pulse arrival time information every 32~ns to a Crate Trigger Processor board (CTP), which is in the same 
crate. The 3-bit pulse arrival time allows the trigger to know the pulse timing at 4~ns resolution. 
Contrary to the readout path 
described in Sec.~\ref{sec:fadc}, this energy is a pedestal-subtracted time-over-threshold sum with 
programmable offsets and minimum threshold discriminator for each channel. With input from all 
FADC channels, i.e. one half of the ECal, the CTP performs cluster finding and calculates cluster 
energy and timing information. The 3x3 fixed-window, highly parallel, FPGA-based cluster 
algorithm simultaneously searches for up to 125 clusters with energy sum larger than 
the programmable energy threshold set to about 270~MeV. This high threshold didn't hurt the trigger rate 
rate studies for the HPS Test run since only clusters with high energies were studied but for HPS this 
threshold will need to be lower.  Crystals in the 
fixed-window are included in the sum if the leading edge of the pulse occurred within a 32~ns time 
window to take into account clock skew and jitter throughout the system.
The CTP only accepts clusters with the locally highest energy 3x3 window to deal with overlapping and 
very large clusters. The Sub-System Processor board (SSP) receives the clusters from the top and bottom half CTP 
at a maximum rate of 250~MHz and searches for pairs of clusters in an 8~ns wide coincidence window. The 
SSP sends triggers to the trigger supervisor (TS), which generates all the necessary signals and 
controls the entire DAQ system readout through the trigger interface units installed in every crate that 
participate in the readout process.

@@

During most of the HPS Test Run, the trigger system 
required only a single cluster in either the top or bottom ECal halves resulting in rates below 2~kHz. However, 
the trigger system was tested to rates above 100~kHz by lowering thresholds.

@@

The SVT DAQ is based on the Reconfigurable Cluster Element (RCE) and cluster 
interconnect concept developed at SLAC as generic building blocks for DAQ systems. 
The RCE is a generic computational building block, housed on a separate daughter card called 
Data Processing Module (DPM), that is realized on an ATCA front board called the Cluster On Board 
(COB), as shown in Figure~\ref{fig:cob}.
 \begin{figure}[]
\begin{center}
\includegraphics[width=6cm]{svt_daq_module}
\caption{\small The SVT DAQ COB board with four data processing daughter cards (DPMs) visible on the left side.
\label{fig:cob}}
\end{center}
\end{figure}
The first generation RCE used in the HPS Test Run consisted of a Virtex~5 FPGA with 1~GB of DDR3 RAM. 
A schematic overview of the system is shown in Figure~\ref{fig:svtdaq}. 
 \begin{figure}[]
\begin{center}
\includegraphics[width=8cm]{svt-daq-sketch}
\caption{\small Block diagram overview of the SVT DAQ.
\label{fig:svtdaq}}
\end{center}
\end{figure}
The analog outputs of up to 12 SVT half-modules (60 APV25 ASICs) are digitized on the 
Rear-Transition-Module (RTM), a custom board on the back side of the ATCA crate, interfacing the 
HPS-specific readout to the generic DAQ components on the COB. A preamplifier converts the APV25 ASIC 
differential current output to a different voltage output scaled to the sensitive range of a 14-bit ADC 
operating at the system clock of 41.667~MHz. The RTM is organized 
into four sections with each section supporting three SVT half-module hybrids (15 APV25 ASICs). The 
RTM also includes a 4-channel fiber-optic module and supporting logic which is used to interface 
to the JLab trigger system supervisor. Each section of the RTM is input to a DPM which apply 
thresholds for data reduction and organizes the sample data into UDP datagrams. The DPM also hosts 
an I$^{2}$C controller used to configure and monitor the APV25 ASICs. A single ATCA crate with two 
COB cards was used, one supporting four DPMs and one supporting three DPMs and one DPM that is 
configured as the trigger and data transmission module. The two COB cards and their DPMs are 
interconnected with a 10~Gb/s switch card~\cite{Larsen:2011zb} which also hosts two 1~Gb/s Ethernet 
interfaces to the external SVT DAQ PC.

@@

In order to minimize cable length for the analog APV25 ASIC output signal the ATCA crate was located 
approximately 1~m from the beam line, next to the cable vacuum feed-throughs.   
Before shielding with lead-blankets and borated polyethylene was arranged, we observed two failures of normally reliable ATCA 
crate power supplies, time-correlated to beam instabilities.

@@

After a trigger is received, the amplitude of every APV25 ASIC is sampled and digitized in the six consecutive time 
bins associated with the trigger time. A data reduction algorithm is applied 
requiring three out of six samples to be above two times the noise level and that the third sample is 
larger than the second or that the fourth sample is larger than the third.
The typical, pedestal subtracted, pulse shape obtained is shown in Figure~\ref{fig:pulse_shape}. 
As the figure demonstrates,  the SVT was well timed-in to the trigger with the rise of the pulse at the 3rd 
sampling point. 
\begin{figure}[]
\begin{center}
\includegraphics[width=7cm]{run1351_110513_samples_L1_top.pdf}
\caption{\small Accumulation of six pedestal-subtracted samples from individual SVT channels associated 
with hits on tracks.
\label{fig:pulse_shape}}
\end{center}
\end{figure}
In order to find the time, $t_0$, and amplitude of each hit, the six samples from each channel are fitted 
to an ideal $CR-RC$ function. Note that in the HPS Test Run the APV25 ASICs were operating with a 50~ns 
shaping time. These hits are passed through a simple clustering algorithm which forms clusters by 
grouping adjacent strips with the position of a cluster on the sensor determined by the 
amplitude-weighted mean.
With a linear gain up to $\approx 3$~MIPs, the cluster charge for hits associated with a track follow 
the characteristic Landau shape, see Figure~\ref{fig:cluster_pulse}.  
\begin{figure}[]
\begin{center}
\includegraphics[width=7cm]{mip_top_layer_2.png}
\caption{\small The cluster charge distribution for hits associated with a track follow the characteristic Landau shape. 
\label{fig:cluster_pulse}}
\end{center}
\end{figure}
A noise level between $1.1-1.5\times 10^{3}$ electrons was established through multiple calibration 
runs giving a $S/N$ of $21-25$, in line with the requirement for HPS. 
Radioactive source tests were used to 
provide the absolute charge normalization.
After clustering hits on a sensor, the hit time for each cluster is computed 
as the amplitude-weighted average of the individually fitted $t_0$ on each channel. The 
$t_0$ resolution is studied by comparing the cluster hit time with the average of all cluster hit times on 
the track shown in Figure~\ref{fig:tracktime}. 
After correcting for offsets from each sensor (time-of-flight and clock phase) and accounting for the 
correlation between the $t_0$ and track time,  the extracted $t_0$ resolution is 2.6~ns. This is 
somewhat worse than the approximately 2~ns resolution expected for S/N$=$25 which we attribute to the 
true pulse shape differing from our idealized fit function which will be improved in the 
future~\cite{Friedl2007385}. Reducing 
the APV25 ASIC pulse shaping time to 35~ns will also improve time resolution. 
These results show that HPS can operate with the six sample readout mode of the APV25 ASIC and 
achieve time resolution adequate for pileup rejection during electron running in HPS. 
\begin{figure}[]
\begin{center}
\includegraphics[width=7cm]{run1351_time_residual_top_layer2.pdf}
\caption{\small The residual of individual cluster times with the average of all clusters on the track. 
\label{fig:tracktime}}
\end{center}
\end{figure}

@@

\subsection{Trigger Rate Comparisons}
Trigger rates observed in the HPS Test Run are dominated by \ee{} pairs produced in the 
converter. In simulated events, the rate of triggers depend on the modeling of the pairs' angular 
distribution and the subsequent multiple Coulomb scattering in the converter. Rates from different 
converter 
thicknesses are used to study the varying multiple Coulomb scattering contribution (pair production 
angular distribution is constant).  These rates are compared with those predicted by the standard Geant4~\cite{Agostinelli2003250} approximation for multiple 
scattering and the more accurate model used by EGS5~\cite{egs5}. 
%\begin{center}
\begin{table}
{\small
\begin{tabular}{|l|c|c|c|}
\hline
\bf Converter (\%~$X_0$) & \bf 1.60 & \bf 0.45 &	\bf 0.18 \\
\hline
EGS5 &	1162 $\pm$ 112 &	255 $\pm$ 28 &	94 $\pm$ 17	\\
\hline
Geant4 & 2633 $\pm$ 250 & 	371 $\pm$ 38 &	114 $\pm$ 18 \\
\hline
Observed 	& 1064 $\pm$ 2 & 196 $\pm$ 1 &	92 $\pm$ 1 \\						
\hline
\end{tabular}
\caption{ Observed and predicted event rate (in Hz) normalized to 90~nA for three different converter 
thicknesses. The uncertainty on the prediction includes systematic uncertainties from ECal alignment, background normalization, beam current normalization and limited statistics in the simulation.
\label{results}}
}
\end{table}
%\end{center}
Restricting to a well calibrated region of the ECal and to clusters with energy above the trigger turn-on, 
we see agreement with the rates predicted by the \egs{} simulation program after subtracting the 
``no converter'' background , see Table~\ref{results}.
This gives further confidence that the dominant source of background occupancy for HPS, multiple 
Coulomb scattered beam electrons, is well 
described.
%~\cite{Attwood:2005zz,Shen:1978ha,Gottschalk1993467} .
